---
title: "An Exploration Into Craft Beers"
author: "Lester Pi"
date: "4/4/2017"
output:
  html_document: default
  pdf_document: default
---
```{r Setup, echo=FALSE, warning=FALSE, include=FALSE}
library('XML')
library('RCurl')
library('plyr')
library(Quandl)
library("quantmod")
require("xts")
library('tseries')
library(RJSONIO)
require(MASS)
require(knitr)
require(rmarkdown)
library(forecast)
library(sqldf)
library(hashmap)
library(ggplot2)
library(maps)
library(tcltk)
```


```{r global_options, echo=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(fig.path='Figs/')
```

```{r set Width, echo=FALSE}
opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
opts_chunk$set(dev = 'pdf')
```


#Introduction

I was browsing Kaggle one day and saw a data set about craft beers and breweries in the US. This piqued my interest, so I decided to download it and take a look at it. The main areas of interest within the data set that I wanted to take a look at is the Alcohol By Volume (ABV) and the state that each brewery is in. Note that DC is included as a state.

Some questions I wanted to look into is if different properties of the beer has statistical significance in the ABV and whether the economic and social environment in a state factors into the ABV in beers from that state. For my analysis, I converted ABV to percentage points to make the results nicer. 

You might ask, "Wouldn't alcohol consumption be a better indicator for economic and social environment?". Well, I considered that, but the answer seemed a bit too obvious and this data can tell a more interesting story.

I acknowledge that the data is not perfect and purposely stuck with the Kaggle data. The data used is the most up to date I could find at the time. For the purpose of exploration of curiousity, I will not focus too much on data quality.

\newpage
#A Look Into Craft Beers Statistics


```{r, echo=FALSE, warning=FALSE}
#source https://www.kaggle.com/nickhould/craft-cans
#source https://wallethub.com/edu/happiest-states/6959/
#source FRED
setwd("C:/cygwin64/home/Lester/beer")


#trim leading and trailing whitespace from string s
trimWhitespace = function (s){
  s=sub("^\\s+", "", s)
  s=sub("\\s+$", "", s)
  return(s)
}

#match a regular expression, return true or false
regexpMatch=function(pattern,string){
  regexp = regexpr(pattern,string)
  if(regexp[1]==-1){
    return(FALSE)
  }
  else{
    return(TRUE)
  }
}

#initial load and transform and sanity checks
beers = read.csv("beers.csv", header = TRUE)
breweries = read.csv("breweries.csv", header = TRUE)
colnames(breweries)[1] = "brewery_id"

beers_sql = sqldf('SELECT * FROM beers WHERE brewery_id=1')
breweries_sql = sqldf('SELECT * FROM breweries WHERE brewery_id=1')




#data grabbed from fred website
#used most recent data available, not all were upt to date
#removed sd, no beer production
state_h = read.csv("state_income_per_capita.csv")
state_h$state = as.character(state_h$state)
state_h$state = trimWhitespace(state_h$state)
state_h$per_capita_income = as.numeric(state_h$per_capita_income)
state_h$h = as.numeric(state_h$h)
state_h$emotional_h = as.numeric(state_h$emotional_h)
state_h$workplace_h = as.numeric(state_h$workplace_h)
state_h$env_h = as.numeric(state_h$env_h)

state_h = state_h[order(state_h$state),]

beers_full = merge(beers,breweries,by="brewery_id")
#style to char for regex extraction later
beers_full$style = as.character(beers_full$style)
beers_full$style = trimWhitespace(beers_full$style)
beers_full$abvPoints = beers_full$abv*100
```

To start things off, here is a peek into the craft beer data set after merging the beers and brewery data sets.

```{r,echo=FALSE}
head(beers_full)
```

Some of the columns of interest to me, other than state and abv, are the ibu and style. The ibu column represents the International Bittering Units (IBU), which is a measurement of how bitter a beer is. The bitterness of a beer is most commonly attributed to the amount of hops used during the brewing process. This leads me to look at the beer style column, which tells us what kind of beer it is.

I ran a regression of ABV on IBU to see the effect IBU has on ABV.

```{r,echo=FALSE}

beers_no_NA = subset(beers_full,!is.na(beers_full$abv))

no_ibu_na = na.omit(beers_no_NA)
ibu_reg = lm(abvPoints~ibu,data=no_ibu_na)
summary(ibu_reg)
plot(no_ibu_na$ibu,no_ibu_na$abvPoints,col="skyblue",main="ABV on IBU",ylab="ABV",xlab="IBU")

lines(no_ibu_na$ibu,ibu_reg$fitted.values,type='l',col="red")

```

Its easy to see there is an obvious relationship between IBU and ABV. The coefficient on ABV is strong and statistically significant and the R Squared isn't too bad either. IBU, however, does not have a causal relationship on ABV. ABV is usually due to how many sugars are fermented during the making of the beer.


```{r,echo=FALSE}

#abv_state=lm(abvPoints~state,data=beers_full)
#summary(abv_state)

beers_orderby_state = sqldf('SELECT * FROM beers_full ORDER BY state')

#create and organize data



beers_full_stateAsString = beers_no_NA
beers_full_stateAsString$state = as.character(beers_full_stateAsString$state)
beers_full_stateAsString$state = trimWhitespace(beers_full_stateAsString$state)
beers_full_stateAsString = sqldf('SELECT * FROM beers_full_stateAsString ORDER BY state')
#beers_full_stateAsString$state

#sanity check
check_AL = sqldf("SELECT * FROM beers_full_stateAsString WHERE state = 'AL'")
#check_AL
mean_AL = mean(check_AL$abvPoints)
#mean_AL
#summary(abv_state)

#use this data set
beers_usethis = beers_full_stateAsString


```

#The Myth of IPAs Being Stronger

It is common to think that IPAs are stronger than other beers. I believe this is can be explained by looking into the bitterness relationship. 

I was able to group the different types of IPAs into one IPA group through a regex match.

A side note, some good information about why IPAs have a higher ABV can be found here: http://www.titletownbrewing.com/bgk-why-are-hoppy-beers-so-strong

```{r,echo=FALSE}

ipas = beers_usethis
ipas$ipa = 0
i = 1
while(i<=length(ipas$ipa)){
  if(regexpMatch("IPA",ipas$style[i])){
    ipas$ipa[i]=1
  }
  i=i+1
}
ipas_reg1=lm(beers_usethis$abvPoints~ipas$ipa)
summary(ipas_reg1)

```

We can see that it looks as if being an IPA does contribute to a higher ABV. However, I would like to see if "being an IPA" actually has an effect, or is it the effect of the higher IBU associated with IPAs.

```{r,echo=FALSE}
ipas_reg2=lm(beers_usethis$abvPoints~ipas$ipa+ipas$ibu)
summary(ipas_reg2)
```

After controlling for IBUs in the regression, it looks as if the previous regression is a bit misleading. Being an IPA actually makes a beer have a lower ABV and an IPA's higher ABV is corelated to the IBUs, but not caused. 

So, what does this all mean?

It means the other beer styles which would have the same IBU as an IPA would tend to also have a higher ABV. Since the alcohol content from beer comes from the fermentation of sugars from the malt, it makes sense that other beers with more malt would have a higher ABV after controlling for IBUs. So, "being an IPA" does not make a beer stronger. An equally bitter beer that is not an IPA will tend to have more sugars in the brewing process to help balance the bitterness. The IPA style doesn't try to hide the bitterness with extra sugars, hence leading to an explanation of our results.

#Breakdowning Down Beers by States

Can craft beers tell us anything about a state's economic and social environment? Probably not, but it would be interesting if it could, so lets see just for the sake of curiousity.

The data set labelled state_h contains state data on the average ABV, personal income per capita, and four different types of happiness ratings.

Note that some states have a very small sample size when it comes to craft beers and breweries, but I will continue this part of the analysis without excluding them anyways. I will average each states' ABV of craft beers so each state has an equal weight and assume this average holds since there is a lack of data and to keep things interesting.

First, I'll see if the per capita income has any effect.

```{r,echo=FALSE}
#create hashmap k=state v=avg abv points


#blank hmap template
key="state"
value="0"
hmap_state_abv = hashmap(key,value)
hmap_state_abv$erase(key)

#fill with state avg abv points data
i = 1
end = length(beers_usethis$state)
while(i <= end){
  if(i==1){
    prev = beers_usethis$state[i]
    tempVals = c()
  }
  state = beers_usethis$state[i]
  curVal = beers_usethis$abvPoints[i]
  if(state!=prev || i == end){
    if(i==end){
      tempVals=c(tempVals,curVal)
    }
    key = prev
    average = mean(tempVals,na.rm=TRUE)
    value = average
    if(hmap_state_abv$has_key(prev)){
      print(paste("key already exists:",key,"at",as.character(i),sep=" "))
    }
    else{
      hmap_state_abv$insert(key,value)
    }
    tempVals = c(curVal)
  }
  if(state==prev){
    tempVals = c(tempVals, curVal)
  }
  
  prev=state
  i = i+1
}

#sanity check
#hmap_state_abv
#hmap_state_abv$find("AL")
#check_WY = sqldf("SELECT abvPoints FROM beers_usethis WHERE state = 'WY'")
#check_WY
#mean_WY = mean(check_WY$abvPoints)
#mean_WY
#summary(abv_state)

#hmap_state_abv



#iterate through states
i = 1
#abv_mean ordered by state abreviation
abv_mean = c()
while(i <= length(state_h[,2])){
  abv_mean=c(abv_mean,as.numeric(hmap_state_abv$find(state_h[,2][i])))
  
  i=i+1
}
state_h$abv=abv_mean

#regress suff
reg = lm(abv_mean~state_h$per_capita_income)
summary(reg)
plot(state_h$per_capita_income,abv_mean,
     ylab="ABV Percentage Points",xlab="Per Capita Income",
     main = "ABV of Craft Beer on Per Capita Income",
     col="skyblue")
lines(state_h$per_capita_income,reg$fitted.values,col="red")

```

Per capita income has no effect on ABV. The plot shows how the data points are all over the place and a linear fit cannot explain it.

Let's examine happiness. There is an overall happiness total score (higher = more happy), emotional & physical well-being ranking (lower = more happy), workplace environment ranking (lower = more happy), and community & environment ranking (lower = more happy).

```{r,echo=FALSE}
happy_reg = lm(state_h$abv~state_h$h+state_h$emotional_h+state_h$workplace_h+state_h$env_h)
summary(happy_reg)
```

We can see there is nothing significant here, but workplace happiness has the strongest significance.

Let's break this down even more by looking at just the happiness score, since it is an aggregate score of overall happiness.

```{r,echo=FALSE}
#just happiness
h_reg = lm(state_h$abv~state_h$h)
summary(h_reg)
```

It looks to be slightly significant, but I want to examine this further by removing irrelevant variables. I removed the intermediate regression results for easier readability.

```{r,echo=FALSE}
#remove insignificant vars
#happy_reg2 = lm(state_h$abv~state_h$h+state_h$emotional_h+state_h$workplace_h)
#summary(happy_reg2)
#happy_reg3 = lm(state_h$abv~state_h$h+state_h$workplace_h)
#summary(happy_reg3)
workplace_h_reg = lm(state_h$abv~state_h$workplace_h)
summary(workplace_h_reg)

```

This is interesting. It turns out that workplace happiness rating was making the overall happiness score look more significant than it really is. It is also interesting that the more unhappy the working environment is in a state, the higher the ABV in their craft beers.

Here is a plot to visualize the results.

```{r,echo=FALSE}

#explore data
#state_h=state_h[order(state_h$workplace_h),]
#state_h
plot(state_h$workplace_h,state_h$abv,
     xlab="Workplace Happiness Rank",ylab="ABV Points",
     main = "ABV points on Workplace Happiness",
     col="skyblue")
lines(state_h$workplace_h,workplace_h_reg$fitted,col="red")
```

There is an obvious trend, even though it doesn't have a tight fit.

These results are extremely thought provoking results, perhaps more so than the relationship between alcohol consumption and workplace happiness. 

Why? 

Because these results can possibly tell us about consumer preferance and a consumer and producer relationship. It is possible that the breweries know their audience and cater to their local crowd. Many craft breweries only supply in their immediate areas. It is easy to see how a worse workplace environment can cause (yes, causation not correlation) a preferance for more alcohol. Now, we can say that there is a correlation between workplace happiness and ABV in craft beers that could point to the producer-consumer relationship.

It could also point to a story about who is crafting these beers. Consider this. A career paper-pusher with an incompetant manager stuck in a dead-end job decides to quit their job and pursue their dream in starting a craft brewery. Let's call this person "J". As a result of the poor work environment, J has developed a preferance for higher alochol consumption. When J sets off to create some top of the line craft beers, J brews what J likes, which turns out to be higher ABV beers that stemmed from J's preferance for higher alochol consumption.

Can we find anything more out of consumer preferances? Since we found IPAs have a higher ABV, does this mean the states with a higher ABV preferance is due to a preferance for IPAs? If the assumption that the producer-consumer relationship holds, by finding the ratio of IPAs produced may give us some insight on this question.

```{r,echo=FALSE}





#find out number of IPAs by state
#iterate through beer_usethis and count occurances of IPA
#regress abv on ipa count
key="state"
value=0
hmap_ipa = hashmap(key,value)
hmap_ipa$erase(key)



#find ipas
i = 1
while(i<= length(beers_usethis[,1])){
  if(regexpMatch("IPA",beers_usethis$style[i])){
    #does not exist, create
    if(!hmap_ipa$has_key(beers_usethis$state[i])){
      hmap_ipa$insert(beers_usethis$state[i],1)
    }
    #exists, incr
    else{
      temp = hmap_ipa$find(beers_usethis$state[i])
      temp=temp+1
      hmap_ipa$insert(beers_usethis$state[i],temp)
    }
  }
  i=i+1
}



#find not ipas
key="state"
value=0
hmap_not_ipa = hashmap(key,value)
hmap_not_ipa$erase(key)

i = 1
while(i<= length(beers_usethis[,1])){
  if(!regexpMatch("IPA",beers_usethis$style[i])){
    #does not exist, create
    if(!hmap_not_ipa$has_key(beers_usethis$state[i])){
      hmap_not_ipa$insert(beers_usethis$state[i],1)
    }
    #exists, incr
    else{
      temp = hmap_not_ipa$find(beers_usethis$state[i])
      temp=temp+1
      hmap_not_ipa$insert(beers_usethis$state[i],temp)
    }
  }
  i=i+1
}

#find ratio of ipa to non ipa
ipa_df = data.frame(state=character(),ipa=numeric(),not_ipa=numeric(),stringsAsFactors=FALSE)
i = 1
while(i <= length(state_h$state)){
  state = state_h$state[i]
  ipa = hmap_ipa$find(state)
  not_ipa = hmap_not_ipa$find(state)
  row=c(state,ipa,not_ipa)
  ipa_df[nrow(ipa_df) + 1, ] = row
  i=i+1
}
ipa_df[is.na(ipa_df)] = 0
ipa_df$ipa=as.numeric(ipa_df$ipa)
ipa_df$not_ipa=as.numeric(ipa_df$not_ipa)
ipa_df$ratio=ipa_df$ipa/ipa_df$not_ipa

#ipa_df

ipa_reg = lm(state_h$abv~ipa_df$ratio)
summary(ipa_reg)

```

It turns out the IPA ratio does not have an effect. This only strengthens our original results on the effects of workplace happiness.

\newpage
#Some Cool Maps


```{r,echo=FALSE}
#plot map

map = state_h
map$state_full=as.character(map$state_full)
map$state_full=trimWhitespace(map$state_full)
map$state_full=tolower(map$state_full)
map$region = map$state_full

all_states <- map_data("state")
map=merge(map,all_states,by="region")
#map
map <- map[map$region!="district of columbia",]
map=map[order(map$order),]



p <- ggplot()
p <- p + geom_polygon(data=map, aes(x=long, y=lat, group = group, fill=map$abv),colour="white"
) + scale_fill_continuous(low = "thistle2", high = "darkred", guide="colorbar")
P1 <- p + theme_bw()  + labs(fill = "ABV points" 
                             ,title = "United States Craft Brew ABV", x="", y="")
#P1 + scale_y_continuous(breaks=c()) + scale_x_continuous(breaks=c()) + theme(panel.border =  element_blank())

P1
```


```{r,echo=FALSE}
p2 <- ggplot()
p2 <- p2 + geom_polygon(data=map, aes(x=long, y=lat, group = group, fill=map$workplace_h),colour="white"
) + scale_fill_continuous(low = "thistle2", high = "darkred", guide="colorbar")
P2 <- p2 + theme_bw()  + labs(fill = "Happiness" 
                             ,title = "United States Workplace Happiness (1 = happiest)", x="", y="")
#P2 + scale_y_continuous(breaks=c()) + scale_x_continuous(breaks=c()) + theme(panel.border =  element_blank())

P2


```

\newpage
#Sources

Kaggle Craft Beer Data: https://www.kaggle.com/nickhould/craft-cans

US Happiness Data: https://wallethub.com/edu/happiest-states/6959/

US Economic Data: https://fred.stlouisfed.org/
